\morus: Biases in the Stream

### 2018-05-10 morusAC_03_MiniMorus.tex
propose introduction (poorly written)
homogenized references
minor consistency fix: MORUS

\section{Symmetric Linear Combinations and \cipher{MiniMORUS}}
\label{sec:introminimorus}

@@

Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
\begin{definition}
Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
\[
S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
\]
is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
\[
\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
\]
\end{definition}
\emph{Example.} The following linear combination is symmetric for \cipher{MORUS640}, i.e. $w = 32$:
\begin{equation}
S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
\label{eq:symmetric}
\end{equation}

@@

\subsection{\cipher{MiniMORUS}}

In fact, we can go further and consider a reduced version of \cipher{MORUS} where each register contains a single word instead of four. The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see Figure~\ref{fig:minimorus}. We call these reduced versions \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}, for \cipher{MORUS640} and \cipher{MORUS1280} respectively. Since registers in \cipher{MiniMORUS} contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.

Because the trail we are going to build is relatively complex, we will first describe it on \cipher{MiniMORUS}. We will then extend it to the full \cipher{MORUS} via the previous symmetry.

\begin{figure}[h]
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
    \printstate
  \end{tikzpicture}%}}}
  \caption{\cipher{MiniMORUS} state update function.}
  \label{fig:minimorus}
\end{figure}

>>>>

%!TEX root = morusAC.tex

\section{Symmetric Linear Combinations and \MiniMORUS}
\label{sec/introminimorus}

@@

Since we want to build a linear trail, we are interested in linear combinations of bits within a register.
\begin{definition}
Recall that $w$ denotes the word size in bits, and $4w$ is the size of a register. A linear combination of the form:
\[
S^t_{i,j(0)} \oplus S^t_{i,j(1)} \oplus \dots \oplus S^t_{i,j(k)}
\]
is said to be \emph{symmetric} iff the set of bits $S^t_{i,j(0)}, \dots, S^t_{i,j(k)}$ is left invariant by a circular shift by $w$ bits; that is, iff:
\[
\{j(i) : i\leq k\} = \{j(i) + w \text{\rm{} mod } 4w : i\leq k\}.
\]
\end{definition}
\emph{Example.} The following linear combination is symmetric for \MORUS[640], i.e. $w = 32$:
\begin{equation}
S^t_{0,0} \oplus S^t_{0,32} \oplus S^t_{0,64} \oplus S^t_{0,96}.
\label{eq:symmetric}
\end{equation}

@@

\subsection{\MiniMORUS}

In fact, we can go further and consider a reduced version of \MORUS where each register contains a single word instead of four. The \StateUpdate{} function is unchanged, except for the fact that register-wise rotations are removed: see \Cref{fig:minimorus}. We call these reduced versions \MiniMORUS[640] and \MiniMORUS[1280], for \MORUS[640] and \MORUS[1280] respectively. Since registers in \MiniMORUS contain a single word, word-wise and register-wise rotations are the same operation; for simplicity we write $\lll$ for word-wise rotations.

Because the trail we are going to build is relatively complex, we will first describe it on \MiniMORUS. We will then extend it to the full \MORUS via the previous symmetry.

\begin{figure}[h]
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \tikzsetnextfilename{minimorus_plain}
  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
    \printstate
  \end{tikzpicture}%}}}
  \caption{\MiniMORUS state update function.}
  \label{fig:minimorus}
\end{figure}
